---
title: "NUTRISHIELD_Study_II_FAMES_HM"
author: "Parth Doshi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = "png",
                      dpi = 300,
                      echo = TRUE,
                      cache = FALSE)
```

## Exploratory Data Analysis for NUTRISHIELD Study II FAMES Human Milk Sample

## Required library

```{r message=FALSE, warning=FALSE}
# install required library
library(devtools)
#devtools::install_github("sfirke/janitor")

# load Required libraries
library(tidyverse) # meta package of all tidyverse packages
library(janitor) # 
library(ggplot2)
library(corrplot)
library(caret)
library(stringr)
library(FactoMineR)
library(ggfortify)
library(ggcorrplot)
library(factoextra)
library(reshape2)
library(Boruta)
library(moments)
library(skimr)
library(car)
```

```{r}
# Set Working Directory
setwd("C:/Users/Parth Doshi/Dropbox/Nutrishield_Study_II_Project (ParthD thesis)/R-script/EDA_for_Human_Mother_Milk")
```

## Data importing & Cleaning 
```{r}
# Load the FAMES.HM.Data from the CSV file
FAMES.HM.Data <- read.csv("C:/Users/Parth Doshi/Dropbox/Nutrishield_Study_II_Project (ParthD thesis)/Study2-Data/NUTRISHIELD_Study_II_FAMES_HM.csv", sep = ";", skip = 3) 
str(FAMES.HM.Data)

FAMES.HM.Data <- FAMES.HM.Data %>%
  # Select relevant columns by excluding "String" and columns containing "NUTRISHIELD" or "X0"
  select(-String, -contains(c("NUTRISHIELD", "X0"))) %>%
  # Remove columns with any missing values
  select_if(~ !any(is.na(.))) %>%
  # Transpose the data to convert it from wide to long format
  t() %>%
  as.data.frame() %>%
  # Set the first row as column names
  row_to_names(1) %>%
  # Convert variables to numeric (except for the "Class" column)
  mutate_if(names(.) != "Class", as.numeric)


colnames(FAMES.HM.Data) <- gsub("\\s*\\(.*?\\)\\s*", "", colnames(FAMES.HM.Data))

# Remove any leading/trailing whitespaces from the column names
colnames(FAMES.HM.Data) <- trimws(colnames(FAMES.HM.Data))

# Add a new column "Samples" with row names
FAMES.HM.Data <- rownames_to_column(FAMES.HM.Data, var = "Samples")

# summary Stats
skimmed_FAMES <- skim(FAMES.HM.Data)
skimmed_FAMES

# Save the clean data to a new CSV file
write.csv(FAMES.HM.Data, "C:/Users/Parth Doshi/Dropbox/Nutrishield_Study_II_Project (ParthD thesis)/Study2-clean-Data/NUTRISHIELD_Study_II_FAMES_HM.csv", row.names = FALSE)

```

### Remove feature with constant value
```{r}
check.constant.or.few.unique <- function(data) {
 result <- vector("character", length = ncol(data))
 
 for (i in 1:ncol(data)) {
   column <- data[[i]]
   
   if (length(unique(column)) == 1) {
     result[i] <- "Constant value"
   } else if (length(unique(column)) <= 10) {
     result[i] <- "Few unique values"
   } else {
     result[i] <- "No pattern"
   }
 }
 pattern_columns <- names(data)[result != "No pattern"]
 return(pattern_columns)
}
results <- check.constant.or.few.unique(FAMES.HM.Data)
print(results)
```

### Remove the variable with constant value and few unique values 

```{r}
FAMES.HM.Data <- FAMES.HM.Data %>% 
  select(-Undecanoate,-Tridecanoate,-Tricosanoate,-Heneicosanoate,-Docosanoate,-`cis-10-Pentadecenoic`,-`cis-11,14,17-Eicosatrienoic`)

# Checking Skewness 
skewness(FAMES.HM.Data[3:31])
```

## Normalization of Data

```{r}
# Normalization of  data
Preprocess <- preProcess(FAMES.HM.Data, method = "BoxCox")
# standardize the preprocessed data
FAMES.HM.Data.Normal <- predict(Preprocess,FAMES.HM.Data)

skewness(FAMES.HM.Data.Normal[3:31])


skimmed_FAMES.normal <- skim(FAMES.HM.Data.Normal)
skimmed_FAMES.normal



# Converting it into long 
FAMES.HM.Data.Normal.long <- gather(FAMES.HM.Data.Normal[3:15])
FAMES.HM.Data.Normal.long.subset <- gather(FAMES.HM.Data.Normal[16:31])

# box plot after distribution
FAMES.HM.Data.Normal.boxplot.A <- ggplot(FAMES.HM.Data.Normal.long, aes(x = value, y = key, fill = key))+
  geom_boxplot(outlier.colour = "red", outlier.shape = 1) +
  labs(title = "Scaled FAMES Data Boxplot",)
print(FAMES.HM.Data.Normal.boxplot.A)


# box plot after distribution
FAMES.HM.Data.Normal.boxplot.B <- ggplot(FAMES.HM.Data.Normal.long.subset, aes(x = value, y = key, fill = key))+
  geom_boxplot(outlier.colour = "red", outlier.shape = 1) +
  labs(title = "Scaled FAMES Data Boxplot",)
print(FAMES.HM.Data.Normal.boxplot.B)
```

### Box plot of data distribution of each group (Post,Pre,PI,TI)

```{r}
FAMES.HM.Data.Normal1 <- subset(FAMES.HM.Data.Normal[1:10])
FAMES.HM.Data.Normal2 <- cbind(FAMES.HM.Data.Normal[1:2],subset(FAMES.HM.Data.Normal[11:20]))
FAMES.HM.Data.Normal3 <- cbind(FAMES.HM.Data.Normal[1:2],subset(FAMES.HM.Data.Normal[21:31]))

# Box plot for 
boxplot <- function(data){
data %>%
    gather(key = "feature", value = "value", -Samples, -Class) %>%
    ggplot(aes(x = Class, y = value, fill = Class)) +
    geom_boxplot() +
    facet_wrap(~ feature, scales = "free") +
    labs(x = "Group", y = "Value")
}

FAMES.HM.Boxplot.1 <- boxplot(FAMES.HM.Data.Normal1)
print(FAMES.HM.Boxplot.1)
FAMES.HM.Boxplot.2 <- boxplot(FAMES.HM.Data.Normal2)
print(FAMES.HM.Boxplot.2)
FAMES.HM.Boxplot.3 <- boxplot(FAMES.HM.Data.Normal3)
print(FAMES.HM.Boxplot.3)
```
## Correlation plot 
```{r}
# Calculate p-values for each correlation coefficient
p.mat <- cor_pmat(FAMES.HM.Data.Normal[3:31], method = "pearson")

# Calculate correlation matrix using Pearson correlation
correlationMatrix <- cor(FAMES.HM.Data.Normal[3:31], method = "pearson")

# Visualize the correlation matrix using ggcorrplot
ggcorrplot(
  correlationMatrix,
  hc.order = TRUE,   # Hierarchical clustering for reordering variables
  type = "lower",    # Show only the lower triangle of the correlation matrix
  lab = FALSE,        # Show labels for variables
  p.mat = p.mat      # Overlay p-values on the plot
)
```

### statistical tests to find relationship between categorical variable and a continuous variable.

```{r}

# Perform Shapiro-Wilk test for each numeric variable to check normalitty
for (i in 3:ncol(FAMES.HM.Data.Normal)) {
  result <- shapiro.test(FAMES.HM.Data.Normal[, i])
  print(result)
}

# Get the names of the continuous variables
continuous.vars <-c("Hexanoate","Octanoate","Decanoate","Laurate","Myristate",    "Myristoleate","Pentadecanoate","Palmitate","Palmitoleate",                 "Heptadecanoate","`cis-10-Heptadecenoic`","Stearate","Oleic","Elaidic","Linoleic",                        "Linolelaidic","`gamma-Linolenic`","Linolenic","Eicosanoic","`cis-11-Eicosenoic`",                  "`cis-11,14-Eicosadienoic`","`cis-8,11,14-Eicosatrienoic`","Arachidonic",                           "`cis-5,8,11,14,17-Eicosapentaenoic`","`Erucic acid`","`cis-13,16-Docosadienoic`",          "`cis-4,7,10,13,16,19-Docosahexaenoic`","Lignocerate","`Nervonic acid`")        

categorical.var <- "Class"

FAMES.HM.Data.Normal$Class <- factor(FAMES.HM.Data.Normal$Class,labels = c("Pre","Post","TI","PI"))

# perform levene's test to see if all group has equal variance
for (vars in continuous.vars) {
  formula <- as.formula(paste(vars, "~", categorical.var))
  var_test <- leveneTest(formula, data = FAMES.HM.Data.Normal)
  cat("Variable:", vars, "\n")
  print(var_test)
  cat("\n")
}

# Perform Kruskal-Wallis test for each variable
for (vars in continuous.vars) {
  formula <- as.formula(paste(vars, "~", categorical.var))
  kruskal_model <- kruskal.test(formula, data = FAMES.HM.Data.Normal)
  cat("Variable:", vars, "\n")
  print(kruskal_model)
  cat("\n")
}

# Loop over each continuous variable
for (var in continuous.vars) {
# Create a formula specifying the relationship between the categorical variable and the current continuous variable
  formula <- as.formula(paste(var, "~ Class"))
  
  # Fit the Anova
  AOV.model <- aov(formula, data = FAMES.HM.Data.Normal)
  tukey_TEST<- TukeyHSD(AOV.model)
  # Print the variable name and the ANOVA table
  cat("Variable:", var, "\n")
  print(summary(AOV.model))
  print(tukey_TEST)
  cat("\n")
}
```

#### Regression test

```{r}
for (var in continuous.vars) {
  # Create a formula specifying the relationship between the continuous variable and the categorical variable
  formula <- as.formula(paste(var, "~ Class"))
  
  # Fit the linear regression model
  regression.model <- lm(formula, data = FAMES.HM.Data.Normal)
  
  # Print the regression coefficients and summary
  cat("Variable:", var, "\n")
  print(coef(regression.model))
  cat("\n")
  print(summary(regression.model))
  cat("\n")
}

```

##### Principle component Analysis
```{r}
FAMES.HM.PCA <- prcomp(FAMES.HM.Data.Normal[3:31], scale. = TRUE, center = TRUE)

#Visualization of PCA
fviz_pca_ind(FAMES.HM.PCA,
             geom = "point",
             habillage = FAMES.HM.Data.Normal$Class,
             palette = c("blue", "red","green","yellow"),
             addEllipses = TRUE,
             ellipse.type="confidence",
             ggtheme = theme_bw(),
             title = "PCA plot for HM Fatty Acids")
#Scree plot 
fviz_eig(FAMES.HM.PCA, 
         addlabels = TRUE, 
         ylim = c(0, 70),
         main="Scree Plot Fatty Acids")

#Graph for variable
fviz_pca_var(FAMES.HM.PCA, col.var = "red")
```

# feature selection
```{r}
FAMES.HM.Data.Normal$Class <- factor(FAMES.HM.Data.Normal$Class,labels = c("Pre","Post","TI","PI"))

# define the control using a random forest selection function
control <- rfeControl(functions=rfFuncs, method="cv", number=10)
# run the RFE algorithm
results <- rfe(FAMES.HM.Data.Normal[,3:29], FAMES.HM.Data.Normal[,2], sizes=c(1:29), rfeControl=control)
# summarize the results
print(results)
# list the chosen features
predictors(results)


library(Boruta)
# Decide if a variable is important or not using Boruta
boruta_output <- Boruta(Class ~ ., data=na.omit(FAMES.HM.Data.Normal), doTrace=2)  # perform Boruta search

boruta_signif <- names(boruta_output$finalDecision[boruta_output$finalDecision %in% c("Confirmed", "Tentative")])  # collect Confirmed and Tentative variables

plot(boruta_output, cex.axis=.7, las=2, xlab="", main="Variable Importance")  # plot variable
```