---
title: "NUTRISHIELD_Study_II_QTrap_Targeted_Urine_EDA"
author: "Parth Doshi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = "png",
                      dpi = 300,
                      echo = TRUE,
                      cache = TRUE)
```

## Exploratory Data Analysis for NUTRISHIELD Study II QTrap Targeted Urine Sample

## Required library

```{r message=FALSE, warning=FALSE}
# install required library
library(devtools)
#devtools::install_github("sfirke/janitor")

# load Required libraries
library(tidyverse) # meta package of all tidyverse packages
library(janitor)
library(ggplot2)
library(caret)
library(ggcorrplot)
library(FactoMineR)
library(ggfortify)
library(factoextra)
library(reshape2)
library(ggbiplot)
library(stringr)
library(corrplot)
library(moments)
library(skimr)
library(car)
```

## Data importing

```{r}
# Set Working Directory
setwd("C:/Users/Parth Doshi/Dropbox/Nutrishield_Study_II_Project (ParthD thesis)/R-script/EDA_for_Human_Mother_Milk")

# load Data set
Targeted.biomarkers.data <- read.csv("C:/Users/Parth Doshi/Dropbox/Nutrishield_Study_II_Project (ParthD thesis)/Study2-Data/NUTRISHIELD_Study_II_QTrap_Targeted_Urine.csv", sep = ";", skip = 3)
str(Targeted.biomarkers.data)
```

## Data Cleaning 

```{r}
# Data cleaning for Targeted Biomarkers Data

# Select the columns containing "IU" or "MU" in their names
Targeted.biomarkers.data <- Targeted.biomarkers.data %>%
  select(Sample, contains(c("IU", "MU"))) %>%

  # Transpose the data to convert it from wide to long format
  t() %>%
  as.data.frame() %>%

  # Set the first row as column names
  row_to_names(1) %>%
  # Convert variables to double (numeric) except for the "Class" column
  mutate_if(names(.) != "Class", as.double) %>%
  # Rename the rownames column as "Samples"
  rownames_to_column(var = "Samples")

# Remove the "(umol/g creat)" part from the column names using regex
colnames(Targeted.biomarkers.data) <- gsub("\\s*\\(umol\\/g creat\\)", "", colnames(Targeted.biomarkers.data))

# Display the Summary statistic
skimmed_target <-  skim_to_list(Targeted.biomarkers.data)
skimmed_target

# Save the clean data to a CSV file
write.csv(Targeted.biomarkers.data, "C:/Users/Parth Doshi/Dropbox/Nutrishield_Study_II_Project (ParthD thesis)/Study2-clean-Data/NUTRISHIELD_Study_II_QTrap_Targeted.csv", row.names = FALSE)

```
### Remove feature with constant value
```{r}
# Function to check if features has Constant value or very few unique values
check.constant.or.few.unique <- function(data) {
 result <- vector("character", length = ncol(data))
 
 for (i in 1:ncol(data)) {
   column <- data[[i]]
   
   if (length(unique(column)) == 1) {
     result[i] <- "Constant value"
   } else if (length(unique(column)) <= 50) {
     result[i] <- "Few unique values"
   } else {
     result[i] <- "No pattern"
   }
 }
 pattern_columns <- names(data)[result != "No pattern"]
 return(pattern_columns)
}
results <- check.constant.or.few.unique(Targeted.biomarkers.data)
print(results)
```
##### Remove the variable with constant value and few unique values 
```{r}
Targeted.biomarkers.data <- Targeted.biomarkers.data %>% 
  select(-`3-IPA`,-Daidzein,-Equol,-Glycitein,-Genistein)
```
### Data Preprocessing 

```{r}
# Data Distribution
# Gather columns 3 to 14 into a long format
Targeted.biomarkers.data.long.A <- gather(Targeted.biomarkers.data[3:14])

# Gather columns 15 to 24 into a long format
Targeted.biomarkers.data.long.B <- gather(Targeted.biomarkers.data[15:24])

# Generate a box plot for Targeted.biomarkers.data.long.A
ggplot(Targeted.biomarkers.data.long.A, aes(x = value, y = key, fill = key)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 1) +
  labs(title = "Data Distribution box plot before normalization")

# Generate a box plot for Targeted.biomarkers.data.long.B
ggplot(Targeted.biomarkers.data.long.B, aes(x = value, y = key, fill = key)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 1) +
  labs(title = "Data Distribution box plot before normalization")

# Calculate skewness for columns 3 to 24
skewness(Targeted.biomarkers.data[3:24])

```

### Transfromation of the data using Log10 Tranformation 
From the skewness you can see data is right skewed therefore using log10 transformation to reduce  skewness

```{r warning=FALSE}
# log 10 transformation
# Perform log10 transformation on columns 3 to 24
Targeted.biomarkers.data.Normal <- as.data.frame(log10(Targeted.biomarkers.data[3:24]))

# Replace infinite values with 0
for (col in names(Targeted.biomarkers.data.Normal)) {
  Targeted.biomarkers.data.Normal[[col]][!is.finite(Targeted.biomarkers.data.Normal[[col]])] <- 0
}

# Combine the transformed data with the first two columns from the original data
Targeted.biomarkers.data.Normal <- cbind(Targeted.biomarkers.data[1:2], Targeted.biomarkers.data.Normal)

# Generate summary statistics for the transformed data
skimmed_target_normal <- skim_to_list(Targeted.biomarkers.data.Normal)
skimmed_target_normal

# Calculate skewness after the transformation
skewness(Targeted.biomarkers.data.Normal[3:24])

# Box plot after the transformation
Targeted.biomarkers.data.Normal.long.A <- gather(Targeted.biomarkers.data.Normal[3:14])
Targeted.biomarkers.data.Normal.long.B <- gather(Targeted.biomarkers.data.Normal[15:24])

# Generate a box plot for Targeted.biomarkers.data.Normal.long.A
ggplot(Targeted.biomarkers.data.Normal.long.A, aes(x = value, y = key, fill = key)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 1) +
  labs(title = "Data Distribution box plot before normalization") +
  xlim(-1, 5)

# Generate a box plot for Targeted.biomarkers.data.Normal.long.B
ggplot(Targeted.biomarkers.data.Normal.long.B, aes(x = value, y = key, fill = key)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 1) +
  labs(title = "Data Distribution box plot after normalization") +
  xlim(-1, 5)

```

```{r}
# Box plot of data distribution of each group (Mother , PI , TI)
Targeted.biomarkers.data.Normal %>%
  gather(key = "feature", value = "value", -Samples, -Class) %>%
  ggplot(aes(x = Class, y = value, fill = Class)) +
  geom_boxplot() +
  facet_wrap(~ feature, scales = "free") +
  labs(x = "Group", y = "Value")
```

```{r}
# Removing the variable which has large  no of outliers  
Targeted.biomarkers.data.Normal <- Targeted.biomarkers.data.Normal %>% 
  select(-Quercetin, -Kaempferol, -`1-Methylhistidine`)
```

```{r}
# Principle component Analysis
Targeted.biomarkers.Data.Normalise.pca <- prcomp(Targeted.biomarkers.data.Normal[3:22], scale. = TRUE, center = TRUE)

#Visualization of PCA
fviz_pca_ind(Targeted.biomarkers.Data.Normalise.pca,
             geom = "point",
             habillage = Targeted.biomarkers.data.Normal$Class,
             palette = c("blue", "red","yellow"),
             addEllipses = TRUE,
             ellipse.type="confidence",
             ggtheme = theme_bw(),
             title = "PCA plot for Qtrap targeted biomarkrs")

#Scree plot 
fviz_eig(Targeted.biomarkers.Data.Normalise.pca, 
         addlabels = TRUE, 
         ylim = c(0, 70),
         main="Scree Plot Qtrap targeted Biomarkers")

#Graph for variable
fviz_pca_var(Targeted.biomarkers.Data.Normalise.pca, col.var = "red")
 
# biplot
biplot = ggbiplot(pcobj = Targeted.biomarkers.Data.Normalise.pca,
                  choices = c(1,2),
                  scale = 0,
                  varname.size = 2.5,
                  varname.abbrev = FALSE,  # Abbreviate variable names (TR
                  var.axes = TRUE,      # Remove variable vectors (TRUE)
                  circle = FALSE,        # Add unit variance circle (TRUE
                  ellipse = FALSE, groups = Targeted.biomarkers.data.Normal$Class) # Adding ellipses
print(biplot)
```
## Correlation & Covariance
```{r}
# Calculate p-values for each correlation coefficient
p.mat <- cor_pmat(Targeted.biomarkers.data.Normal[3:22], method = "pearson")

# Calculate correlation matrix using Pearson correlation
correlationMatrix <- cor(Targeted.biomarkers.data.Normal[3:22], method = "pearson")

# Visualize the correlation matrix using ggcorrplot
ggcorrplot(
  correlationMatrix,
  hc.order = TRUE,   # Hierarchical clustering for reordering variables
  type = "lower",    # Show only the lower triangle of the correlation matrix
  lab = FALSE,        # Show labels for variables
  p.mat = p.mat      # Overlay p-values on the plot
)

# Calculate the covariance matrix using Pearson correlation
cov_df <- cov(Targeted.biomarkers.data.Normal[sapply(Targeted.biomarkers.data.Normal, is.numeric)], method = "pearson")

# Convert covariance matrix to correlation matrix
cov_df <- cov2cor(cov_df)

# Reshape the correlation matrix to long format
cov_df_long <- melt(cov_df)

# Create the covariance plot using ggplot2
plot <- ggplot(cov_df_long, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient(high = "red", low = "white") +
  geom_text(aes(label = paste0(round(value * 100, 2))), color = "black", size = 3) +
  labs(title = "Covariance Heatmap",
       x = "Variable 1",
       y = "Variable 2")

# Print the covariance plot
print(plot)

```
### Test of Homogeneity of Variance And ANOVA
```{r}

# Perform Shapiro-Wilk test for each numeric variable to check normality
for (i in 3:ncol(Targeted.biomarkers.data.Normal)) {
  variable_name <- names(Targeted.biomarkers.data.Normal)[i]
  result <- shapiro.test(Targeted.biomarkers.data.Normal[, i])
  cat("Variable:", variable_name, "\n")
  print(result)
  cat("\n")
}


# Define the Continuous Variables
continuous.vars <- c("Phenylpropionylglycine", "Taurine", "Citrulline", "Galactitol", "Isovalerylglycine", "Isobutyrylglycine", "TMAO", "Anserine", "`3-Methylhistidine`", "`L-Kynurenine`", "`L-Tyrosine`", "`O-DMA`", "Phloretin", "Hesperetin", "`Proline betaine`", "`Ferullic Acid Sulfate`", "`Hippuric Acid`", "`3-IAA`")

categorical.var <- "Class"
Targeted.biomarkers.data.Normal$Class <- factor(Targeted.biomarkers.data.Normal$Class)

# Perform Levene's test to see if all groups have equal variance
for (vars in continuous.vars) {
  formula <- as.formula(paste(vars, "~", categorical.var))
  var_test <- leveneTest(formula, data = Targeted.biomarkers.data.Normal)
  cat("Variable:", vars, "\n")
  print(var_test)
  cat("\n")
}

# Perform Kruskal-Wallis test for each variable
for (vars in continuous.vars) {
  formula <- as.formula(paste(vars, "~", categorical.var))
  kruskal_model <- kruskal.test(formula, data = Targeted.biomarkers.data.Normal)
  cat("Variable:", vars, "\n")
  print(kruskal_model)
  cat("\n")
}

# Perform ANOVA and display summary for each model
for (vars in continuous.vars) {
  formula <- as.formula(paste(vars, "~", categorical.var))
  aov_model <- aov(formula, data = Targeted.biomarkers.data.Normal)
  tukey_test <- TukeyHSD(aov_model)
  cat("Variable:", vars, "\n")
  print(summary(aov_model))
  print(tukey_test)
  cat("\n")
}

```

### Regression Test
```{r}
# perform regression 
for (Vars in continuous.vars) {
  formula <- as.formula(paste(Vars, "~", categorical.var))
  Lm.model <-lm(formula, data = Targeted.biomarkers.data.Normal)
  cat("Variable:",Vars, "\n")
  print(summary(Lm.model))
  cat("\n")
}
```


## Feature selection
#### Feature Selection Using Recursive feature elimination Method
```{r}
Targeted.biomarkers.data.Normal$Class <- factor(Targeted.biomarkers.data.Normal$Class, labels = c("Mother","TI","PI"))
# define the control using a random forest selection function
control <- rfeControl(functions=rfFuncs, method="cv", number=10)
# run the RFE algorithm
results <- rfe(Targeted.biomarkers.data.Normal[,3:22], Targeted.biomarkers.data.Normal[,2], sizes=c(1:22), rfeControl=control)
# summarize the results
print(results)
# list the chosen features
predictors(results)


library(Boruta)
# Decide if a variable is important or not using Boruta
boruta_output <- Boruta(Class ~ ., data=na.omit(Targeted.biomarkers.data.Normal), doTrace=2)  # perform Boruta search

boruta_signif <- names(boruta_output$finalDecision[boruta_output$finalDecision %in% c("Confirmed", "Tentative")])  # collect Confirmed and Tentative variables

# plot variable
plot(boruta_output, cex.axis=.7, las=2, xlab="", main="Variable Importance")  
```