---
title: "NUTRISHIELD_Study_II_MIRIS_HM_EDA"
author: "Parth Doshi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = "png",
                      dpi = 300,
                      echo = TRUE,
                      cache = FALSE)
```

## Exploratory Data Analysis for NUTRISHIELD Study II MIRIS Human Milk Sample

## Required library

```{r message=FALSE, warning=FALSE}
# install required library
library(devtools)
devtools::install_github("sfirke/janitor")
#install.packages("DescTools")
#install.packages("skimr")
# load Required libraries
library(tidyverse) # meta package of all tidyverse packages
library(janitor) # 
library(ggplot2)
library(corrplot)
library(ggcorrplot)
library(caret)
library(FactoMineR)
library(ggfortify)
library(factoextra)
library(moments)
library(reshape2)
library(ggbiplot)
library(skimr)
library(car)
```

## Data importing

```{r}
# Set Working Directory
setwd("C:/Users/Parth Doshi/Dropbox/Nutrishield_Study_II_Project (ParthD thesis)/R-script/EDA_for_Human_Mother_Milk")

#load Data
MIRIS.HM.Data <- read.csv("C:/Users/Parth Doshi/Dropbox/Nutrishield_Study_II_Project (ParthD thesis)/Study2-Data/NUTRISHIELD_Study_II_MIRIS_HM.csv",sep = ",", skip = 2)
str(MIRIS.HM.Data)


MIRIS.HM.Data <-  MIRIS.HM.Data %>%
  # Finding specific pattern to select the sample from data set
  select(-c(1:7,9:12)) %>%
  select_if(~ !any(is.na(.))) %>%
  t() %>%
  as.data.frame() %>%
  row_to_names(1) %>%
  mutate_all(as.numeric)
 
     
MIRIS.HM.Data <- rownames_to_column(MIRIS.HM.Data,var = "Samples")

# assigning new names to the columns of the data frame
colnames(MIRIS.HM.Data) <- c("Samples","Fat","Crude_Protein","Carbohydates","TS","Energy","True_Protein")

write.csv(MIRIS.HM.Data,"C:/Users/Parth Doshi/Dropbox/Nutrishield_Study_II_Project (ParthD thesis)/Study2-clean-Data/NUTRISHIELD_Study_II_MIRIS_HM.csv")
```

```{r}
skimmed_MIRIS <- skim(MIRIS.HM.Data)
skimmed_MIRIS

# check for Skewness
print("skewness Score")
skewness(MIRIS.HM.Data[2:7])
```

## Data Prepossessing

```{r warning=FALSE}

# using caret lib to preprocess data
Normalization <- preProcess(MIRIS.HM.Data, method = c("BoxCox"),na.remove = TRUE )

# standardize the preprocessed data
MIRIS.HM.Data.scale <- predict(Normalization,MIRIS.HM.Data)

# Skewness Score After Normalization
skewness(MIRIS.HM.Data.scale[2:7])

# Summary Stats after Normalization
skimmed_MIRIS_scale <- skim_to_list(MIRIS.HM.Data.scale)
skimmed_MIRIS_scale
```
## Comparative anaylsis 

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Assigning Class to Variable  according to information from sample ID

MIRIS.HM.Data.new <- MIRIS.HM.Data.scale %>% 
  mutate(Class = case_when(grepl("^POST", Samples) ~ "post",
    grepl("^PRE", Samples) ~ "pre",
    grepl("^P", Samples) ~ "PI",
    grepl("^T", Samples) ~ "TI",
    TRUE ~ "other"
  ))

attach(MIRIS.HM.Data)

# Boxplot Function 
boxplot <- function(data){
data %>%
    gather(key = "feature", value = "value",-Class, -Samples) %>% 
    ggplot(aes(x = Class, y = value, fill = Class)) +
    geom_boxplot() +
    facet_wrap(~ feature, scales = "free") +
    labs(x = "Group", y = "Value")
}

Milk.Boxplot <- boxplot(MIRIS.HM.Data.new)
print(Milk.Boxplot)
```

##Exploratory Data Analysis

## correlation 

```{r}
# Calculate p-values for each correlation coefficient
p.mat <- cor_pmat(MIRIS.HM.Data.scale[2:7], method = "pearson")

# Calculate correlation matrix using Pearson correlation
correlationMatrix <- cor(MIRIS.HM.Data.scale[2:7], method = "pearson")

# Visualize the correlation matrix using ggcorrplot
ggcorrplot(
  correlationMatrix,
  hc.order = TRUE,   # Hierarchical clustering for reordering variables
  type = "lower",    # Show only the lower triangle of the correlation matrix
  lab = TRUE,        # Show labels for variables
  p.mat = p.mat      # Overlay p-values on the plot
)
```
### Covariance 

```{r}
# Calculate covariance matrix
cov_matrix <- cov(MIRIS.HM.Data.scale[2:7], method = "pearson")

cov_matrix <- cov2cor(cov_matrix)
cov_df_long <- melt(cov_matrix)


# Create the covariance plot using ggplot2
plot <- ggplot(cov_df_long, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient(high = "red", low = "white") +
  geom_text(aes(label = paste0(round(value * 100, 2))), color = "black", size = 3) +
  labs(title = "Covariance Heatmap",
       x = "Variable 1",
       y = "Variable 2")


print(plot)
```

# PCA 
```{r}
MIRIS.HM.PCA <- prcomp(MIRIS.HM.Data.new[2:7], scale = TRUE)

#Visualization of PCA
fviz_pca_ind(MIRIS.HM.PCA,
             geom = "point",
             habillage = MIRIS.HM.Data.new$Class,
             palette = c("blue", "red","green","yellow"),
             addEllipses = TRUE,
             ellipse.type="confidence",
             ggtheme = theme_bw(),
             title = "PCA plot for HM Fatty Acids")
#Scree plot 
fviz_eig(MIRIS.HM.PCA, 
         addlabels = TRUE, 
         ylim = c(0, 70),
         main="Scree Plot Fatty Acids")

# Variable plot
fviz_pca_var(MIRIS.HM.PCA, col.var = "red")


```

### stastical test to identfy is  significance difference between variables 

```{r}

MIRIS.HM.Data.new$Class <-  factor(MIRIS.HM.Data.new$Class)

# Perform Leven's test for each continuous variable
for (i in 2:7) {
  # Create the formula
  formula <- as.formula(paste(names(MIRIS.HM.Data.new)[i], "~ Class"))
  
  # Perform Levene's test
  var_test <- leveneTest(formula, data = MIRIS.HM.Data.new)
  
  # Print the variable name and the test results
  cat("Variable:", names(MIRIS.HM.Data.new)[i], "\n")
  print(var_test)
  cat("\n")
}



# Perform ANOVA and display summary for each model
for (i in 2:7) {
  # Create the formula
  formula <- as.formula(paste(names(MIRIS.HM.Data.new)[i], "~ Class"))
  # Perform Anova
  aov_model <- aov(formula, data = MIRIS.HM.Data.new)
  # Perform Tukey Test
  tukey_test <- TukeyHSD(aov_model)
  
  
  # Print Results
  cat("Variable:", names(MIRIS.HM.Data.new)[i], "\n")
  print(summary(aov_model))
  print(tukey_test)
  cat("\n")
}
```

### Feature selection 

```{r}
set.seed(100)

ctrl <- rfeControl(functions = rfFuncs,
                   method = "repeatedcv",
                   repeats = 10,
                   verbose = FALSE)

rfe_Profile <- rfe(x=MIRIS.HM.Data.new[, 2:7], y=MIRIS.HM.Data.new$Class,
                 sizes = c(1:8),
                 rfeControl = ctrl)

rfe_Profile
```
### Regression Model 

```{r}
# Split the data into training and test sets (80% training, 20% test)
Train_Index <- createDataPartition(
  y = MIRIS.HM.Data.scale$TS,
  ## the outcome data are needed
  p = .70,
  ## The percentage of data in the
  ## training set
  list = FALSE
)

train_Data <- MIRIS.HM.Data[ Train_Index, -1]
test_Data  <- MIRIS.HM.Data[-Train_Index, -1]
```

# linear regression model

```{r}
fitControl <- trainControl(method = "repeatedcv",   
                           number = 10,     # number of folds
                           repeats = 10)    # repeated ten times

# Train the model using Linear Regression and predict on the training data itself.
lm_model <-  train(TS ~ ., data=train_Data, method='lm', trControl = fitControl)

# predict on test data
lm_model_fitted <- predict(lm_model, test_Data)
lm_model
```
### SVM-radial

```{r}
# Train the model using Linear SVM and predict on the Test data.
Svm_model <-  train(TS ~ ., data=train_Data, method='svmRadial', trControl = fitControl)

# predict on test data
SVM_model_fitted <- predict(Svm_model, test_Data)
Svm_model
```

### KNN-Regression

```{r}
# Train the model using Knn and predict on the test data.
KNN_model <-  train(TS ~ ., data=train_Data, method='knn', trControl = fitControl)

# predict on test data
KNN_model_fitted <- predict(KNN_model, test_Data)
KNN_model
```

### PLS
```{r}
# Train the model using PLS and predict on the Test data.
PLS_model <-  train(TS ~ ., data=train_Data, method='pls', trControl = fitControl, tuneGrid = expand.grid(ncomp = 2))

# predict on train model
PLS_model_fitted <- predict(PLS_model,test_Data)

PLS_model
```

### Random Forest

```{r}
# Train the model using PLS and predict on the Test data.
RF_model <-  train(TS ~ ., data=train_Data, method='rf', trControl = fitControl)

# predict on train model
RF_model_fitted <- predict(RF_model,test_Data)

RF_model
```